AN EFFICIENT LOCALIZED MODEL ORDER REDUCTION FRAMEWORK FOR THE SHAPE OPTIMIZATION OF ADDITIVELY MANUFACTURED LATTICE STRUCTURES\break ECCOMAS CONGRESS 2024

### 2024-08-26 paper/paper.tex
wip: sec 2

The contributions of the present work are given as follows.
First, as references~\cite{Smetana2016Optimal,Taddei2018Localization} do not make use of range approximation via random sampling, a suitable training strategy to construct local approximation spaces for parameterized linear problems via random sampling is discussed.
Herein, the approach given in~\cite{Taddei2018Localization}, identified as a \textit{distributed approximate POD} (see also~\cite{Himpe2018Hierarchical}), is adopted to range approximation via random sampling.
Second, a framework for the shape optimization of lattice structures is proposed.
It combines the aforementioned algorithm to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
Furthermore, the global approximation space is constructed from the local spaces using the generalized finite element method (GFEM)~\cite{Babuska2004Generalized}.

@@

the weak form reads \added[id=pd]{TODO: is it necessary to write Dirichlet lift explicitly?}

@@

Finally, we note that the images of $\bm{T}_{\m}$ are computed, such that they do not contain any translations or rotations (rigid body modes are removed via orthogonal projection), \ie{} they comprise purely \textit{deformational modes}.
Moreover, (global) \textsc{Neumann} or \textsc{Dirichlet} boundary conditions prescribed on $\varSigma^{\m}_{\mathrm{N}}$ or $\varSigma^{\m}_{\mathrm{D}}$, respectively, have not been considered yet.
Therefore, for example, an additional \textsc{Neumann} problem\footnote{The case of inhomogeneous \textsc{Dirichlet} boundary conditions is analogous.} has to be solved to enrich the local spaces.
\begin{align}
	\begin{split}
	\label{eq:neumann_problem}
    - \div{\bm{\sigma}} (\bm{u}(\bm\mu)) &= \bm{0} \quad  \mathrm{in}\;{\varOmega^{\bm\mu}}\subset\varOmega^{\bm\mu}_{\mathrm{gl}}\,,\\
    \bm{\sigma}(\bm{u}(\bm\mu)) \vdot \bm{n} &= \hat{\bm{t}} \quad \mathrm{on}\; \varGamma^{\m}_{\mathrm{N}}:= \partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{N}}\,,\\
    \bm{u}(\bm\mu) &= \bm{0} \quad  \mathrm{on}\;\varGamma^{\m}_{\mathrm{D}}:=\partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{D}}\,,\\
    \bm{u}(\bm\mu) &= \bm{0}\quad\mathrm{on}\;\varGamma^{\m}_{\mathrm{out}}:= \partial\varOmega^{\m} \setminus\partial\varOmega^{\m}_{\mathrm{gl}}\,.
	\end{split}
\end{align}
\Cref{eq:neumann_problem} is solved on the parent domain employing a pull back as above.
Also, for construction of a sufficiently rich local basis it has to be solved for many parameter values $\m$, but a sampling of the boundary values on $\varGamma^{\m}_{\out}$ is not required.
% - setting, global domain = union of non overlapping subdomains = unit cells
% - introduction of coarse grid where each cell corresponds to unit cell subdomain
% - goal of constructing local spaces for only parts of the domain
% - definition of oversampling problem and transfer operator
% these are details of the specific problem we solve later
% - the oversampling problem should be formulated for union of parent unit cells
% - weak formulation is not required here? Should be first introduced in section 2.3?
% - comment on number of archetypes/configurations

@@

\begin{itemize}
    \item assume discretization of T
    \item explain challenge of finding appropriate space
    \item go with idea from Taddei \& Patera, but range approx. instead of TE
\end{itemize}

@@

\added[id=pd]{TODO: Given a discretization of T, we can now discuss how the range can be approximated.}

\added[id=pd]{TODO: check how/when Neumann data is added in case of RRFPOD. This can also be done outside the loop and even after POD of spectral basis.}
\added[id=pd]{TODO: FIX neumann data in the algorithm.}

@@

\subsection{Construction of a global approximation} % (fold)
\label{sub:Construction of a global approximation}

\begin{itemize}
    \item challenge that local spaces are not conforming on the interface
    \item use GFEM as suitable method to construct conforming global approximation
    \item how much detail should be necessary to explain GFEM?
\end{itemize}

@@

\subsection{Online efficiency} % (fold)
\label{sub:Online efficiency}

\begin{itemize}
    \item explain approach based on \textit{super elements}: each unit cell is just a quadrilateral cell with special ansatz functions
    \item challenge: would still need to assemble $K$ of unit cell for each new parameter
    \item therefore, employ EIM to speed up assembly of $K$ (reduced integration)
\end{itemize}

>>>>

The contributions of the present work are given as follows.
First, as references~\cite{Smetana2016Optimal,Taddei2018Localization} do not make use of range approximation via random sampling, a suitable training strategy to construct local approximation spaces for parameterized linear problems via random sampling is discussed.
Herein, the approach given in~\cite{Taddei2018Localization}, identified as a \textit{distributed approximate POD} (see also~\cite{Himpe2018Hierarchical}), is adopted to range approximation via random sampling.
Second, a framework for the shape optimization of lattice structures is proposed.
It combines the aforementioned algorithm to construct local approximation spaces with an auxiliary problem, as in~\cite{Guo2022Learning}, to facilitate geometrical parametrizations of the unit cell and the matrix version~\cite{Negri2015Efficient} of the empirical interpolation method (EIM)~\cite{Barrault2004‘empirical,Chaturantabut2010Nonlinear} to ensure online efficiency of the final reduced order model (ROM).
Furthermore, the global approximation space is constructed from the local spaces using the partition of unity method or generalized finite element method (GFEM)~\cite{BABUSKA1997,Babuska2004Generalized}.

@@

the weak form reads

@@

Finally, we note that the images of $\bm{T}_{\m}$ are computed, such that they do not contain any translations or rotations (rigid body modes are removed via orthogonal projection), \ie{} they comprise purely \textit{deformational modes}.
Moreover, (global) \textsc{Neumann} or \textsc{Dirichlet} boundary conditions prescribed on $\varSigma^{\m}_{\mathrm{N}}$ or $\varSigma^{\m}_{\mathrm{D}}$, respectively, have not been considered yet.
Therefore, for example, an additional \textsc{Neumann} problem\footnote{The case of inhomogeneous \textsc{Dirichlet} boundary conditions is analogous.} has to be solved to enrich the local spaces.
\begin{align}
	\begin{split}
	\label{eq:neumann_problem}
    - \div{\bm{\sigma}} (\bm{u}(\bm\mu)) &= \bm{0} \quad  \mathrm{in}\;{\varOmega^{\bm\mu}}\subset\varOmega^{\bm\mu}_{\mathrm{gl}}\,,\\
    \bm{\sigma}(\bm{u}(\bm\mu)) \vdot \bm{n} &= \hat{\bm{t}} \quad \mathrm{on}\; \varGamma^{\m}_{\mathrm{N}}:= \partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{N}}\,,\\
    \bm{u}(\bm\mu) &= \bm{0} \quad  \mathrm{on}\;\varGamma^{\m}_{\mathrm{D}}:=\partial\varOmega^{\m}\cap\varSigma^{\m}_{\mathrm{D}}\,,\\
    \bm{u}(\bm\mu) &= \bm{0}\quad\mathrm{on}\;\varGamma^{\m}_{\mathrm{out}}:= \partial\varOmega^{\m} \setminus\partial\varOmega^{\m}_{\mathrm{gl}}\,.
	\end{split}
\end{align}
\Cref{eq:neumann_problem} is solved on the parent domain employing the pull back as above.
Also, for construction of a sufficiently rich local basis it has to be solved for many parameter values $\m$, but a sampling of the boundary values on $\varGamma^{\m}_{\out}$ is not required.
% - setting, global domain = union of non overlapping subdomains = unit cells
% - introduction of coarse grid where each cell corresponds to unit cell subdomain
% - goal of constructing local spaces for only parts of the domain
% - definition of oversampling problem and transfer operator
% these are details of the specific problem we solve later
% - the oversampling problem should be formulated for union of parent unit cells
% - weak formulation is not required here? Should be first introduced in section 2.3?
% - comment on number of archetypes/configurations

@@

Assuming a proper FE discretization of $\bm{T}_{\m}$, next the algorithm to approximate the range of $\bm{T}_{\m}$ for any $\m$ is discussed.
Herein, a strategy for the exploration of the parameter space and sampling of the random boundary conditions in the oversampling are the key challenges.
In the present work, the strategy by Taddei \& Patera~\cite{Taddei2018Localization} is adopted in which based on the connection between the POD of images of the transfer operator (for certain boundary conditions) and the singular values of the transfer operator they devise a training strategy which is recognized as a distributed approximate POD (see \cite{Himpe2018Hierarchical}).
For each $\m_j$ in a predefined training set, they compute the singular values of $\bm{T}_{\m_j}$ by resolution of an eigenvalue problem and subsequently compute the POD over the set of all eigenvectors.
As the authors state, the resolution of the eigenvalue problems may be prohibitive in some cases, and therefore we employ range approximation by random sampling instead.

The procedure for the computation of parameter-independent deformational modes is summarized in~\cref{algo:rrfpod}.
First, an empty snapshot set $\bm{S}$ and a training set $\bm{S}_{\mathrm{train}}$ are initialized.
The training set $\bm{S}_{\mathrm{train}}$ has to be sufficiently rich, such that it is a good representation of the parameter space.
\comment[id=pd]{TODO: add reference to Haasdonk (standard sampling)?}
Next, for each parameter value $\m_j\in\bm{S}_{\mathrm{train}}$ the range of $\bm{T}_{\m_j}$ which is now fixed with respect to the parameter value is approximated using Algorithm 1 \textsc{AdaptiveRandomizedRangeApproximation} from~\cite{Buhr2018Randomized} and the set of snapshots is extended by Gram Schmidt orthogonalization (line 5 \& 6).
Note that the parameter value $\m$ and boundary conditions $\bm{g}$ are not varied at the same time.
Finally, the POD is computed over the snapshot set $\bm{S}$ which yields the parameter-independent space $X^n$.
The dimension of the local space $X^n$, \ie{} the number of local basis functions is denoted by $n$.

In case of inhomogeneous \textsc{Neumann} boundary conditions, after termination of~\cref{algo:rrfpod}, the \textsc{Neumann} problem~\cref{eq:neumann_problem} is solved for each $\m_j\in\bm{S}_{\mathrm{train}}$ as well.
Following compression via POD the space $X^n$ is enriched by the additional \textsc{Neumann} modes.

@@

\comment[id=pd]{TODO: explain tolerances to control approximation quality. Refer to HAPOD paper and mention that $\varepsilon^{\ast}$ can be used to compute tolerances at each node of the tree.}
\comment[id=pd]{TODO: define notation for POD.}

@@

\subsection{Construction of a global approximation} % (fold)
\label{sub:Construction of a global approximation}
Given local approximation spaces $X^n_i$ for each target subdomain $\varOmega_i^{\p}$, the construction of a global approximation via the GFEM is discussed.
The difficulty is that the basis functions of the local spaces $X^n_i$ and $X^n_j$ of two neighboring subdomains $\varOmega_i^{\p}$ and $\varOmega_j^{\p}$ are not conforming on the shared interface $\varGamma^{\p}_{ij}=\partial\varOmega_i^{\p}\cap\partial\varOmega_j^{\p}$.
However, given a suitably\footnote{According to Definition 1 of~\cite{BABUSKA1997}.} defined partition of unity $\{{\varphi}_i\}$ the local basis functions can be included in the global FE space by multiplication with the standard FE shape functions ${\varphi}_i$, such that a conforming approximation is obtained.
The global GFEM space is defined as
\begin{equation}
	\label{eq:Xgfem}
	X_{\mathrm{GFEM}}:= \sum_{i} \varphi_iX^n_i = \left\lbrace\sum_i\varphi_i\bm\xi_i \;\big|\; \bm\xi_i\in X^n_i	\right\rbrace\,.
\end{equation}
\comment[id=pd]{TODO: add figure. This explains best what is going on.}
\comment[id=pd]{TODO: equation for partition of unity and final gfem space. Not sure if I can write this like Babuska.}

@@

\subsection{Online efficiency} % (fold)
\label{sub:Online efficiency}
With the global GFEM space, each coarse grid cell (\ie{} unit cell of the lattice) can be viewed as a high order FE.
For example, it is just a quadrilateral cell with special ansatz functions 
\begin{itemize}
    \item explain approach based on \textit{super elements}: each unit cell is just a quadrilateral cell with special ansatz functions
    \item challenge: would still need to assemble $K$ of unit cell for each new parameter
    \item therefore, employ EIM to speed up assembly of $K$ (reduced integration)
\end{itemize}
